Ignition of Fuel Beds by Firebrands

### 2022-05-30 manuscript3.tex
Updates and added model github link

    Fuel beds at risk for ignition by firebrands vary significantly in chemical composition which may result in differences in thermal properties and the chemical species produces during pyrolysis. Works considering ignition of fuel beds with varying concentrations of constituents have found significant variation in ignition characteristics. Cellulose fuel beds have been observed to have a lower ignition threshold when compared to grass and pine needles when put in contact by hot metal particles~\cite{Urban2018}. Interestingly, cellulose undergoes pyrolysis at either higher temperatures or later in the heating process than both lignin and hemicellulose~\cite{Yang2007a, Shotorban2018}. This suggests that the lower ignition temperature of cellulose is not due to pyrolysis occurring at lower temperatures, but to the flammability of pyrolyzates produced.

    One method to evaluate the effects of chemical composition on the flammability of pyrolyzates produced, and thus ignition propensity, is through modeling of the pyrolysis process. Efforts to characterize the pyrolysis of biomass have continually improved the understanding of the chemical processes that occur in both the solid and gas phase of pyrolysis products. Recently, more inclusive pyrolysis models have been developed that improve on the initial single step Arrhenius reactions~\cite{DIBLASI199371} and more closely approximate the reactions occurring during combustion of biomass~\cite{Ranzi2008, Debiagi2015, Dhahak2019} based on the initial chemical composition of the fuels. With the development of models that are applicable to a wide range of chemical compositions and environmental factors the possibility of modelling pyrolysis and ignition of fuel beds is possible. Coupling these models with databases, such as the Bioenergy Feedstock Library~\cite{feedstock} (which contains the chemical composition information for a wide variety of materials that may be vulnerable to firebrand attack during a wildfire), predictions of ignition potential may be possible without data from ignition experiments. While currently available tools, like combustion models and composition databases, may be useful individually for understanding fuel bed ignition and pyrolysis, a framework linking the individual resources and knowledge to provide a general model of fuel bed ignition is lacking. Unfortunately, the current level of knowledge on the ignition threshold of materials of variable composition does not facilitate such a framework.

@@

    The objective of this work is to identify how changes in chemical composition of fuel bed materials effect the probability of flaming ignition. It is anticipated that knowledge gained from this study may be used improve the understanding of fuel bed ignition for different fuels and help build a framework for estimating ignition probabilities of different fuels without the need for extensive laboratory testing.

@@

    The 50\% probability of flaming ignition with respect to the surface temperature of a resistance heater was determined for five different materials in quiescent conditions and at a wind speed of 5.8\si{\meter\per\second}. The fuel bed materials used were Douglas-fir wood, Douglas-fir bark, pine wood, pine bark, wheat straw, and oak wood. These materials were chosen to represent a range of chemical compositions, and are materials that may be subject to firebrand attack in a wildland urban interface fire. Table \ref{tab:composition} shows the estimated chemical composition of the materials based on the Bioenergy Feedstock Library~\cite{feedstock}. To create fuel beds of consistent particle size,  materials were granulated and then sorted such that the particles fit through a screen with openings of 2.1\si{\milli\meter} but not through openings of 0.85\si{\milli\meter}. Once sorted, the materials were oven dried at 103\si{\celsius}.

@@

    Two experimental apparatus were used to conduct ignition experiments. The experimental apparatus used for the tests in quiescent conditions is shown in Figure~\ref{fig:speciesApparatus}. The quiescent apparatus used a lever arm to hold the cartridge heater in a fixed position the duration of the test. The heater was inserted into the fuel bed approximately half of the heater diameter (3\si{\milli\meter}). Tests conducted with a wind speed of \si{\meter\per\second} were performed using the apparatus shown in Figure~\ref{fig:speciesWindTunnel}. This apparatus was operated inside of a wind tunnel where the heater was lowered using an automated lowering device that maintained a constant pressure equivalent to that of a 10\si{\gram} firebrand landing on the fuel bed. The pressure was monitored and adjusted using a PID control system with force measured by a load cell. For all tests the cartridge heater (firebrand surrogate) used was a 50\si{\milli\meter} long 6.35\si{\milli\meter} diameter cartridge heater.

@@

    The temperature of the heater was recorded with a type-K thermocouple for the duration of each test. The temperature of the heater for ranged from 250\si{\celsius} to 750\si{\celsius} and was controlled using PID control logic implemented in LabVIEW. The PID controller kept the heater temperature within 6\% of the setpoint for the duration of the tests. The temperature setpoints of the resistance heater were determined using the three-phase optimal design procedure to efficiently determine the desired ignition probability for the number of tests conducted~\cite{Wu2014}. For tests where flaming ignition did not occur data was recorded for 3000\si{\second} or until the reaction front of the smoldering material reached the edge of the test container. The fixed position of the cartridge heater and the use of the cartridge heater was implemented to maintain consistency between tests of the same material and between test series of different materials. The use of the cartridge heater and the fixed position is not necessarily representative of a burning firebrand landing on the fuel bed but the consistency of surface temperatures, ability to record temperature, and consistent contact with the fuel bed are essential for comparing results across test series and materials. For the quiescent tests flaming ignition was detected with the use of a BPX 65 photodiode sampled at 1 kHz. If the intensity detected rose above a set threshold flaming ignition was said to occur. For the wind tunnel tests flaming ignition was determined from the rising temperature of the cartridge heater in the presence of flame. Visual detection of flames was also used in cohort with the photodiode measurements. The 50\% probability of ignition was determined from a logistic regression on the outcomes of 25 ignition tests conducted for each of the materials for which flaming ignition was observed. The scikit-learn python package~\cite{scikit-learn} was used to perform the logistic regressions used to predict ignition probabilities.

@@

    The thermal conductivity of the porous fuel bed materials was calculated using correlations for porous media from literature~\cite{bergman2011fundamentals}. Equation~\ref{eqn:speciesk} shows the the correlation used where $\epsilon$ is defined as the proportion of volume occupied by air, as is shown in Equation~\ref{eqn:speciesEpsilon}.

@@

    Figure~\ref{fig:composition_plot} shows the averaged cellulose, hemicellulose, and lignin concentrations for each material according to the samples recorded in the Bioenergy Feedstock Library, as shown in Table~\ref{tab:composition}. The marker colors for each material correspond to the calculated 50\% ignition probability for the quiescent tests except for the wheat straw and Douglas-fir bark, which are denoted as 800\si{\celsius} to indicate that the 50\% ignition probability was higher than the temperatures tested. Note that due to scaling inherent to a ternary projection, the proportion of the fuels that is not cellulose, hemicellulose, or lignin, causes a shift in the data. For example, the wheat straw and Douglas-fir are estimated to have the same cellulose content but do not lie on the same iso-line of 44\% cellulose. Nonetheless, the relative proportions of each constituent are retained, and comparisons may still be made. From the data shown in Figure~\ref{fig:composition_plot} a local minimum ignition threshold exists near the concentration of Douglas-fir wood. If the apparent local minimum persists with additional materials, it could be used to rapidly identify materials less conducive to ignition which would be a powerful tool for choosing appropriate materials on and around homes in areas at risk of ember attack. 
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=0.75\columnwidth]{Figures/speciesStudyTernaryDiss.png}
            \caption{Estimated chemical composition of each material tested with the marker color representing the estimated 50\% ignition probability. Note that the materials where flaming ignition did not occur are represented as 800\si{\celsius} to show that the temperature of ignition was not achieved.}
            \label{fig:composition_plot}
        \end{figure}

    The 50\% ignition probability results for the 5.8\si{\meter\per\second} wind speed tests are shown in Table~\ref{tab:composition50temp} along with the results for the 0.1\si{\meter\per\second} tests. From these results, there are three observations of note. First, the increase of wind lowered the threshold for ignition probability for the Douglas-fir wood and Pine wood by approximately 30\%. The decreased threshold for ignition  is attributed to the formation in recirculation zones near the heater. This is consistent with results presented in Chapter~\ref{chap:manuscript2}. Second, in contrast to the Douglas-fir and Pine wood results, an increase in wind resulted in an increase in the ignition threshold for oak wood. Flaming ignition was observed for some tests of oak wood under wind. However, the predicted threshold for ignition was above 750\si{\celsius}. Third, the materials that did not ignite in the quiescent tests were also not observed to ignite in the presence of wind. The trends for Douglas-fir wood and pine wood match those presented in Chapter~\ref{chap:manuscript2} and trends reported in other studies of pine needle ignition~\cite{Wang2017} and eucalyptus bark~\cite{Ellis2011, Ellis2015}. However, the decreasing ignition probability of the oak wood contradicts those observed in the aforementioned studies and this study. A decrease in the ignition in the presence of wind is not unprecedented. An ignition study of various litter layer types ignited by different types of firebrands observed sensitivity to firebrand location. Ignition probability was reduced when the firebrand landed on top of the fuel bed when compared to embedded firebrands~\cite{Plucinski2008}. The decrease in ignition probability was attributed to increased heat loss due to wind. In this study, however, the location of the cartridge heater was consistent across all materials, the energy is not a function of the wind speed, and the flow field around the firebrands is consistent.      
        \begin{table}[hpbt]
            \caption{Heater temperature required for 50\% probability of ignition for the materials and conditions tested.}
            \centering
            \begin{tabular}{crrr}
                % \rowcolor{gray!50}
                Material & 0.1 \si{\meter\per\second}(\si{\celsius}) & 5.8 \si{\meter\per\second} (\si{\celsius}) & $\Delta$T (\si{\celsius})\\
                \hline
                Douglas-fir Wood & 574  & 391 & -185 \\
                Pine Wood        & 634  & 430 & -204\\
                Oak Wood         & 663  & \textgreater750 & 87\\
                Wheat Straw      & \textgreater750 & \textgreater750 & -\\
                Pine Bark        & \textgreater750 & \textgreater750 & -\\
                Douglas-Fir Bark & \textgreater750 & \textgreater750 & -
            \end{tabular}
            \label{tab:composition50temp}
        \end{table}
    Considering the consistent ignition heat source and environmental conditions, the remaining differences that may be driving factors for ignition trends of oak wood are material properties or changes in pyrolyzate ignition due to wind. The most accessible material property to quantify and compare is the bulk density of the materials. However, there is no clear correlation between density and the probability of ignition. Consider first the wheat straw, which did not ignite at either wind speed, had a bulk density between pine wood and Douglas-fir wood, which ignited in both wind conditions. Based on bulk density, wheat straw is anticipated to ignite at a temperature between these two materials. Similarly, oak wood has a bulk density higher than the Douglas-fir bark and pine bark but was observed to ignite in the low wind speed cases. Clearly, the bulk density of the material is not sufficient to differentiate between cases where ignition does or does not occur.

    The thermal diffusivity of the fuel bed materials may provide insight into ignition trends that bulk density does not capture. From the thermal diffusivity values presented in Table~\ref{tab:speciesKValues} the fuel bed materials with the highest thermal diffusivity values (pine bark, Douglas-fir bark, and oak wood in a wind speed of 5.8\si{\meter\per\second}) were not observed to ignite. While there appears to be a correlation between thermal diffusivity and ignition, the wheat straw had a lower thermal diffusivity than both pine wood and Douglas-fir wood but was not observed to ignite in any tests. It is possible that, similar to the chemical composition results, an ideal thermal diffusivity for ignition exists near the thermal diffusivity of Douglas-fir wood and pine wood, however a more likely cause is error from the multiple correlations used to estimate values or differences in measurements across the multiple studies from which properties were collected. Thus, neither the bulk density nor the thermal diffusivity estimates made here are sufficient to explain differences in ignition, leaving differences in chemical composition as the remaining predictor.

@@

\section{Conclusions}
\label{sec:conclusions3}    
    Flaming ignition tests were conducted for fuel beds of six different materials at two different wind speeds. Ignition of fuel beds was induced by a cartridge heater that served as a firebrand surrogate. The temperature of the heater was maintained at a fixed temperature throughout each test, allowing sensitivities of ignition to fuel bed composition to be evaluated. The ignition probability and chemical composition were used to evaluate the influence of the chemical composition of the fuel bed and observe differences in ignition probabilities in the presence of wind and in quiescent conditions. The specific conclusions of this work are as follows:
        \begin{enumerate}
            \item Of the six materials tested, only three were able to be ignited within below the 750\si{\celsius} maximum temperature of the firebrand surrogate. Two of the three materials for which ignition was not observed were bark. The higher ignition threshold of the bark materials is attributed to higher lignin concentrations. The resistance to ignition present in these materials suggest that using materials high in lignin in locations at risk for firebrand attack in wildfires may reduce ignitions and subsequent losses of structures. It is acknowledged that these results are for flaming ignition only, and high lignin materials may pose a significant risk if smoldering or the smoldering to flaming transition occurs.

@@

            \item A direct correlation between the fuel bed composition and the ignition propensity was not observed. This is attributed to the complexity of biomass pyrolysis. However, a local minimum ignition threshold near the composition of Douglas-fir wood is present that warrants further investigation. 
        \end{enumerate}
    The conclusions of this work show that ignition is sensitive to the chemical composition of fuel beds in both quiescent conditions and under wind. An increase in wind was observed to increase ignition probability for some materials; however, this was not universally the case. Materials high in lignin were among the least likely to ignite and more suitable for placement in areas where firebrands are likely to land or accumulate in a WUI environment. Further and more targeted analysis of the differences in pyrolysis products for each material will likely yield further insight. 
    %\cite{MacLean1941}.

>>>>

    Fuel beds at risk for ignition by firebrands vary significantly in chemical composition, which may result in differences in thermal properties and the chemical species produced during pyrolysis. Works considering ignition of fuel beds with varying concentrations of constituents have found significant variation in ignition characteristics. Cellulose fuel beds have been observed to have a lower ignition threshold when compared to grass and pine needles when put in contact with hot metal particles~\cite{Urban2018}. Interestingly, cellulose undergoes pyrolysis at either higher temperatures or later in the heating process than both lignin and hemicellulose~\cite{Yang2007a, Shotorban2018}. This suggests that the lower ignition temperature of cellulose is not due to pyrolysis occurring at lower temperatures but to the flammability of pyrolyzates produced.

    One method to evaluate the effects of chemical composition on the flammability of pyrolyzates produced, and thus ignition propensity, is through modeling of the pyrolysis process. Efforts to characterize the pyrolysis of biomass have continually improved the understanding of the chemical processes that occur in both the solid and gas phase of pyrolysis products. Recently, more inclusive pyrolysis models have been developed that improve on the initial single step Arrhenius reactions~\cite{DIBLASI199371} and more closely approximate the reactions occurring during combustion of biomass~\cite{Ranzi2008, Debiagi2015, Dhahak2019} based on the initial chemical composition of the fuels. With the development of models that are applicable to a wide range of chemical compositions and environmental factors, the possibility of modeling pyrolysis and ignition of fuel beds is possible. Coupling these models with databases, such as the Bioenergy Feedstock Library~\cite{feedstock} (which contains the chemical composition information for a wide variety of materials that may be vulnerable to firebrand attack during a wildfire), predictions of ignition potential may be possible without data from ignition experiments. While currently available tools, like combustion models and composition databases, may be useful individually for understanding fuel bed ignition and pyrolysis, a framework linking the individual resources and knowledge to provide a general model of fuel bed ignition is lacking. Unfortunately, the current level of knowledge on the ignition threshold of materials of variable composition does not facilitate such a framework.

@@

    The objective of this work is to identify how changes in the chemical composition of fuel bed materials affect the probability of flaming ignition. It is anticipated that knowledge gained from this study may be used to improve the understanding of fuel bed ignition for different fuels and help build a framework for estimating ignition probabilities of different fuels without the need for extensive laboratory testing.

@@

    The 50\% probability of flaming ignition with respect to the surface temperature of a resistance heater was determined for five different materials in quiescent conditions and at a wind speed of 5.8\si{\meter\per\second}. The fuel bed materials used were Douglas-fir wood, Douglas-fir bark, pine wood, pine bark, wheat straw, and oak wood. These materials were chosen to represent a range of chemical compositions and are materials that may be subject to firebrand attack in a wildland urban interface fire. Table \ref{tab:composition} shows the estimated chemical composition of the materials based on the Bioenergy Feedstock Library~\cite{feedstock}. To create fuel beds of consistent particle size,  materials were granulated and then sorted such that the particles fit through a screen with openings of 2.1\si{\milli\meter} but not through openings of 0.85\si{\milli\meter}. Once sorted, the materials were oven-dried at 103\si{\celsius}.

@@

    Two experimental apparatus were used to conduct ignition experiments. The experimental apparatus used for the tests in quiescent conditions is shown in Figure~\ref{fig:speciesApparatus}. The quiescent apparatus used a lever arm to hold the cartridge heater in a fixed position for the duration of the test. The heater was inserted into the fuel bed approximately half of the heater diameter (3\si{\milli\meter}). Tests conducted with a wind speed of \si{\meter\per\second} were performed using the apparatus shown in Figure~\ref{fig:speciesWindTunnel}. This apparatus was operated inside of a wind tunnel where the heater was lowered using an automated lowering device that maintained a constant pressure equivalent to that of a 10\si{\gram} firebrand landing on the fuel bed. The pressure was monitored and adjusted using a PID control system with force measured by a load cell. For all tests the cartridge heater (firebrand surrogate) used was a 50\si{\milli\meter} long 6.35\si{\milli\meter} diameter cartridge heater.

@@

    The temperature of the heater was recorded with a type-K thermocouple for the duration of each test. The temperature of the heater ranged from 250\si{\celsius} to 750\si{\celsius} and was controlled using PID control logic implemented in LabVIEW. The PID controller kept the heater temperature within 6\% of the setpoint for the duration of the tests. The temperature setpoints of the resistance heater were determined using the three-phase optimal design procedure to efficiently determine the desired ignition probability for the number of tests conducted~\cite{Wu2014}. For tests where flaming ignition did not occur, data was recorded for 3000\si{\second} or until the reaction front of the smoldering material reached the edge of the test container. The fixed position of the cartridge heater and the use of the cartridge heater was implemented to maintain consistency between tests of the same material and between test series of different materials. The use of the cartridge heater and the fixed position is not necessarily representative of a burning firebrand landing on the fuel bed, but the consistency of surface temperatures, ability to record temperature, and consistent contact with the fuel bed is essential for comparing results across test series and materials. For the quiescent tests, flaming ignition was detected with the use of a BPX 65 photodiode sampled at 1 kHz. If the intensity detected rose above a set threshold, flaming ignition was said to occur. For the wind tunnel tests, flaming ignition was determined from the rising temperature of the cartridge heater in the presence of flame. Visual detection of flames was also used in cohort with the photodiode measurements. The 50\% probability of ignition was determined from a logistic regression on the outcomes of 25 ignition tests conducted for each of the materials for which flaming ignition was observed. The scikit-learn python package~\cite{scikit-learn} was used to perform the logistic regressions used to predict ignition probabilities.

@@

    The thermal conductivity of the porous fuel bed materials was calculated using correlations for porous media from literature~\cite{bergman2011fundamentals}. Equation~\ref{eqn:speciesk} shows the correlation used where $\epsilon$ is defined as the proportion of volume occupied by air, as is shown in Equation~\ref{eqn:speciesEpsilon}.

@@

    Figure~\ref{fig:composition_plot} shows the averaged cellulose, hemicellulose, and lignin concentrations for each material according to the samples recorded in the Bioenergy Feedstock Library, as shown in Table~\ref{tab:composition}. The marker colors for each material correspond to the calculated 50\% ignition probability for the quiescent tests except for the wheat straw and Douglas-fir bark, which are denoted as 800\si{\celsius} to indicate that the 50\% ignition probability was higher than the temperatures tested. Note that due to scaling inherent to a ternary projection, the proportion of the fuels that are not cellulose, hemicellulose, or lignin, causes a shift in the data. For example, the wheat straw and Douglas-fir are estimated to have the same cellulose content but do not lie on the same iso-line of 44\% cellulose. Nonetheless, the relative proportions of each constituent are retained, and comparisons may still be made. From the data shown in Figure~\ref{fig:composition_plot} a local minimum ignition threshold exists near the concentration of Douglas-fir wood. If the apparent local minimum persists with additional materials, it could be used to rapidly identify materials less conducive to ignition, which would be a powerful tool for choosing appropriate materials on and around homes in areas at risk of ember attack. 
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=0.75\columnwidth]{Figures/speciesStudyTernaryDiss.png}
            \caption{Estimated chemical composition of each material tested with the marker color representing the estimated 50\% ignition probability. Note that the materials where flaming ignition did not occur are represented as 800\si{\celsius} to show that the temperature of ignition was not achieved.}
            \label{fig:composition_plot}
        \end{figure}

    The 50\% ignition probability results for the 5.8\si{\meter\per\second} wind speed tests are shown in Table~\ref{tab:composition50temp} along with the results for the 0.1\si{\meter\per\second} tests. From these results, there are three observations of note. First, the increase of wind lowered the threshold for ignition probability for the Douglas-fir wood and Pine wood by approximately 30\%. The decreased threshold for ignition  is attributed to the formation of recirculation zones near the heater. This is consistent with results presented in Chapter~\ref{chap:manuscript2}. Second, in contrast to the Douglas-fir and Pine wood results, an increase in wind resulted in an increase in the ignition threshold for oak wood. Flaming ignition was observed for some tests of oak wood under wind. However, the predicted threshold for ignition was above 750\si{\celsius}. Third, the materials that did not ignite in the quiescent tests were also not observed to ignite in the presence of wind. The trends for Douglas-fir wood and pine wood match those presented in Chapter~\ref{chap:manuscript2} and trends reported in other studies of pine needle ignition~\cite{Wang2017} and eucalyptus bark~\cite{Ellis2011, Ellis2015}. However, the decreasing ignition probability of the oak wood contradicts those observed in the aforementioned studies and this study. A decrease in the ignition in the presence of wind is not unprecedented. An ignition study of various litter layer types ignited by different types of firebrands observed sensitivity to firebrand location. Ignition probability was reduced when the firebrand landed on top of the fuel bed when compared to embedded firebrands~\cite{Plucinski2008}. The decrease in ignition probability was attributed to increased heat loss due to wind. In this study, however, the location of the cartridge heater was consistent across all materials, the energy is not a function of the wind speed, and the flow field around the firebrands is consistent.      
        \begin{table}[hpbt]
            \caption{Heater temperature required for 50\% probability of ignition for the materials and conditions tested.}
            \centering
            \begin{tabular}{crrr}
                % \rowcolor{gray!50}
                Material & 0.1 \si{\meter\per\second}(\si{\celsius}) & 5.8 \si{\meter\per\second} (\si{\celsius}) & $\Delta$T (\si{\celsius})\\
                \hline
                Douglas-fir Wood & 574  & 391 & -185 \\
                Pine Wood        & 634  & 430 & -204\\
                Oak Wood         & 663  & \textgreater750 & 87\\
                Wheat Straw      & \textgreater750 & \textgreater750 & -\\
                Pine Bark        & \textgreater750 & \textgreater750 & -\\
                Douglas-Fir Bark & \textgreater750 & \textgreater750 & -
            \end{tabular}
            \label{tab:composition50temp}
        \end{table}
    Considering the consistent ignition heat source and environmental conditions, the remaining differences that may be driving factors for ignition trends of oak wood are material properties or changes in pyrolyzate ignition due to wind. The most accessible material property to quantify and compare is the bulk density of the materials. However, there is no clear correlation between density and the probability of ignition. Consider first the wheat straw, which did not ignite at either wind speed, had a bulk density between pine wood and Douglas-fir wood, which ignited in both wind conditions. Based on bulk density, wheat straw is anticipated to ignite at a temperature between these two materials. Similarly, oak wood has a bulk density higher than the Douglas-fir bark and pine bark but was observed to ignite in the low wind speed cases. Clearly, the bulk density of the material is not sufficient to differentiate between cases where ignition does or does not occur.

    The thermal diffusivity of the fuel bed materials may provide insight into ignition trends that bulk density does not capture. From the thermal diffusivity values presented in Table~\ref{tab:speciesKValues} the fuel bed materials with the highest thermal diffusivity values (pine bark, Douglas-fir bark, and oak wood in a wind speed of 5.8\si{\meter\per\second}) were not observed to ignite. While there appears to be a correlation between thermal diffusivity and ignition, the wheat straw had a lower thermal diffusivity than both pine wood and Douglas-fir wood but was not observed to ignite in any tests. It is possible that, similar to the chemical composition results, an ideal thermal diffusivity for ignition exists near the thermal diffusivity of Douglas-fir wood and pine wood; however, a more likely cause is the error from the multiple correlations used to estimate values or differences in measurements across the multiple studies from which properties were collected. Thus, neither the bulk density nor the thermal diffusivity estimates made here are sufficient to explain differences in ignition, leaving differences in chemical composition as the remaining predictor.

@@

\section{Conclusions}
\label{sec:conclusions3}    
    Flaming ignition tests were conducted for fuel beds of six different materials at two different wind speeds. Ignition of fuel beds was induced by a cartridge heater that served as a firebrand surrogate. The temperature of the heater was maintained at a fixed temperature throughout each test, allowing sensitivities of ignition to fuel bed composition to be evaluated. The ignition probability and chemical composition were used to evaluate the influence of the chemical composition of the fuel bed and observe differences in ignition probabilities in the presence of wind and in quiescent conditions. The specific conclusions of this work are as follows:
        \begin{enumerate}
            \item Of the six materials tested, only three were able to be ignited below the 750\si{\celsius} maximum temperature of the firebrand surrogate. Two of the three materials for which ignition was not observed were bark. The higher ignition threshold of the bark materials is attributed to higher lignin concentrations. The resistance to ignition present in these materials suggest that using materials high in lignin in locations at risk for firebrand attack in wildfires may reduce ignitions and subsequent losses of structures. It is acknowledged that these results are for flaming ignition only, and high lignin materials may pose a significant risk if smoldering or the smoldering to flaming transition occurs.

@@

            \item A direct correlation between the fuel bed composition and the ignition propensity was not observed. This is attributed to the complexity of biomass pyrolysis. However, a local minimum ignition threshold near the composition of Douglas-fir wood is present that warrants further investigation. 
        \end{enumerate}
    The conclusions of this work show that ignition is sensitive to the chemical composition of fuel beds in both quiescent conditions and under wind. An increase in wind was observed to increase ignition probability for some materials; however, this was not universally the case. Materials high in lignin were among the least likely to ignite and more suitable for placement in areas where firebrands are likely to land or accumulate in a WUI environment. Further and more targeted analysis of the differences in pyrolysis products for each material will likely yield further insight.
